---
title: "SurvivalAnalysis"
author: "Harika Gillela"
date: "5/23/2020"
output: html_document: keep_md: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "README_figs/README-"
)
```


```{r read data}
setwd("/Users/harikagillela/OneDrive - University of South Florida/SDM-git/Survival-Analysis")
#library(textreadr)
df = read.table("LungCancer.rtf", header = FALSE,skip = 24)
dim(df)
names(df)<-c("treatment","celltype","survival_days","status","kscore","months","age","chemo_prior")
df$chemo_prior = as.integer(gsub("[\\},]", "", df$chemo_prior))
df$celltype = factor(df$celltype)
df$celltype = relevel(df$celltype, "1")
str(df)
attach(df)
```


## Descriptive Statistics


```{r DescriptiveStatistics}
table(treatment)
table(status)
summary(survival_days)
hist(age)
table(survival_days)
table(celltype)
```

## Kaplan-Meier


```{r model1}
library(survival)
library(survminer)
# event - status, survival_days -- time 
km1 <- survfit(Surv(survival_days,status)~1)
ggsurvplot(km1, xlab = "Time", ylab = " Survival Probability", data = df)
```

# 1. Kaplan-Meier survival graphs for patients with the test vs standard treatment.

```{r model2}
km2 <- survfit(Surv(survival_days, status) ~ treatment)
ggsurvplot(km2, xlab = "Time", ylab = " Survival Probability", data = df, ggtheme = theme_bw())
```


1. The patients with test treatment type can survive for long time compared to those of standard treatment but the survival probability is very low (0.05 from time period 500 – 1000). 
2. The survival probability for the standard treatment patients is high initially (till 150 days approx.) but later the probability is reduced when compared to the test treatment. 
3. The graph has ended at time period 553 as we don’t have any standard treatment  patient survived beyond that time period.


__Lets check the survival probablity at certain period of time from the model results__


```{r probablity}
summary(km2, times = 365.25)
summary(km2,times = 183)
```


__Survival Probability__
Standard Treatment: 1 year (365 days)	0.0708, 6 months (183 days)	0.2124
Test Treatment ‘2’: 1 year (365 days)	0.1098, 6 months (183 days)	0.2329


## Semi Parametric Models : Cox Proportional Hazrad Model


```{r model3}
cox <- coxph(Surv(survival_days, status) ~ celltype + treatment + age + chemo_prior +kscore + months, method = "breslow")
summary(cox)
```

__Cox proportional Hazard Model__
* The test treatment indicates the higher risk of death when compared the standard treatment. The test treatment increases the hazard rate by 33%..  This model suggests that standard treatment can increase the survival rate when compared to the test treatment.
* Age, Months from Diagnosis will reduce the hazard rate by 1% which is negligible.
* Comparing with the cell type 1, the cell type 2 ,3 and 4 patients have higher risk of death. Among all types of cells, patients with cell type 3 have the highest risk to the death. The cell type 3 increases the hazard rate by 228%. The cell type 2 and 4 increases the hazard rates by 135% and 49% respectively.
* Having prior chemotherapy will increase the risk of death. This will increase the hazard rate by 7%.
* Patients with better Karnofsky score will have low risk of death or more survival time. This will reduce the hazard rate by 4%.


## Exponential, Weibull, and log-logistic parametric model coefficients


```{r model4}
exp <- survreg(Surv(survival_days, status) ~ celltype + treatment + age + chemo_prior + kscore+ months, dist="exponential")


weibull <- survreg(Surv(survival_days, status) ~ celltype + treatment + age + chemo_prior + kscore+ months, dist="weibull")


loglogistic <- survreg(Surv(survival_days, status) ~ celltype + treatment + age + chemo_prior + kscore + months, dist="loglogistic")


library(stargazer)
stargazer(cox, exp, weibull, loglogistic, type="text")
```


__Interpretations__
__Parametric Model: Weibull Model__

* Increase in patient’s age will reduce the risk to death by factor of 1, exp(0.006)
* Months from Diagnosis will increase the risk of death by factor exp(-0.0005) = 0.99
* The cell types 2, 3 and 4 increase the risk of death when compared to cell type 1. The cell type 2 increases the risk of death by 0.43, the cell type 3 increases the death rate by exp(-1.133) = 0.32 and the cell type 4 increases the risk of death by factor exp(-0.398) = 0.67. This tell us that the survival rate is high in case of patients with cancer with cell type 1.
* Prior Chemotherapy will increase the risk to the death by factor 0.99(Weibull and Log logistic). Patients without prior chemotherapy will have 0.99 times better survival rate than the patients with prior chemotherapy.
* The Karnofsky score will reduce the risk of death by factor exp(0.030068) = 1.030(Weibull and Log logistic)  .
* The test treatment type will accelerate the death of the patient by a factor of exp(-0.228) = 1.256”Weibull”.  Standard treatment patients will have 1.256 times better survival than the test treatment patients.


__Parametric Model: Loglogistic Models__


* Increase in patient’s age will reduce the risk to death by factor of 1, exp(0.009)
* Months from Diagnosis will increase the risk of death by factor exp(-0.0005) = 0.99
* The cell types 2, 3 and 4 increase the risk of death when compared to cell type 1. The cell type 2 increases the risk of death by 0.49, the cell type 3 increases the death rate by exp(-0.743) = 0.47 and the cell type 4 reduces the risk of death by factor exp(0.017) = 1.017. 
* Prior Chemotherapy will increase the risk to the death by factor 0.99(Weibull and Log logistic). Patients without prior chemotherapy will have 0.99 times better survival rate than the patients with prior chemotherapy.
* The Karnofsky score will reduce the risk of death by factor exp(0.030068) = 1.030(Weibull and Log logistic)  .
* The test treatment type will accelerate the death of the patient by a factor of exp(-0.08846) = 0.91



